### 2011-04-06 Text/paper.tex
Started writing it as a paper. Need to do the introduction.

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\begin{document}
\begin{abstract}
Recently we investigated the applicability and parameters choice for the two-dimensional case
Bretherton problem using the lattice Boltzmann method (LBM) \cite{kuzmin-binary2d}. This
paper is the continuation of the previous work and is focused on the three-dimensional case to
validate the LBM for simulation of phenomena for rectangular and square capillaries in the
moderate range of Capillary number parameters. 
\end{abstract}

\section{What to examine}
My plan is to investigate preferably the low capillary number regims as far we have interesting
phenomena occuring where: the axisymmetric radiuses for axes and diagonals. As far as it's
computationally very expensive to resolve the film thickness, it will be nice to investigate
whether underresolved films simulations can produce consistent results. The second thing to study
is the velocity pattern which is much easier to examine because the threshold is located for
$Ca\approx 0.6$. The third thing to study is the film thickness dependancy for the range of
capillary numbers. The last thing to study is the applicability of derived by \citet{heil-threedim}
correlations for rectangular channels of different shapes.

\section{Introduction}
While for two-dimensional Bretherton flow the results are usually limited to the circular shape and
only some of them are for the plane-symmetric case \cite{giavedoni-numerical,heil-bretherton},
vast number of experimental results and simulations are available for the three-dimensional case.
It was found that the for rectangular and/or square shaped capillaries there is a transition for
certain capillary number, where the flow changes from the non axisymmetric to axis symmetric case.
Non axisymmetric case is attributed to the case where the axial radius is different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection, see the
example in Fig. \ref{fig:quarter:capillary:capillary01} for a non-axisymmetric case.

The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the flow becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the film
thickness dependency on the capillary number is in Fig. \ref{fig:heil:three:dim}.

>>>>

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\begin{document}
\begin{abstract}
Classical Bretherton problem describes the propagation of gas fingers through liquid media with
thin liquid films between bubbles and the channel walls. The bubble shape and flow patterns are
complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
the applicability and parameters choice for the two-dimensional case Bretherton problem (flow
between plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
\cite{kuzmin-binary2d}. This
work is the continuation of the previous work. It is focused on the three-dimensional capillaries
simulations with rectangular and square crosssections to
validate the binary liquid LBM for  Bretherton phenomena in the moderate range of the capillary
number $0.1\leq Ca \leq 1.0$.  
The flow is driven by a body force,
and
periodic boundary conditions are applied in the streamwise direction. The results show that the
binary liquid model is able to capture a number of phenomena happening in three-dimensional
capillaries , as the existance of the vortex in front of the bubble and the bubble radii dependency
on the capillary number. Therefore, the lattice Boltzmann free energy binary liquid model can be
used to simulate the Bretherton problem with good accuracy.
\end{abstract}

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
narrow channels. Depending on the geometry it was found that the deposited film thickness
is a complicated function of the capillary number $Ca$:
\begin{equation}
\label{capillary:number:definition}
Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma},
\end{equation}
where $U_{\mathrm{bubble}}$ is the propagation bubble velocity, $\mu_{\mathrm{liq}}$ is the
kinematic liquid viscosity and $\gamma$ is the interfacial surface tension between gas and liquid.

For example, the deposited film thickness
is proportional to $Ca^{2/3}$ in the range of small capillary numbers for circular channels
\cite{bretherton,heil-bretherton}. 
The problem of predicting flow patterns and associated mass transfers for the Bretheron-type flows
is of significant interests for chemical industry as it is widely used in chemical monolith
microreactors \cite{kreutzer-pressure-drop}. The large mass transfer can be achieved because of the
large interfacial area and small diffusion lengths \cite{cerro-bubble-train}. The heat transfer is
also increased in comparison with one phase flows \cite{fukugata-levelset}. While it
is possible to calculate the flow analytically for small capillary numbers \cite{bretherton}, it's
not possible to extrapolate it to the wider range of capillary and Reynolds numbers used in the
chemical
industry. Thus, the desire of consistent numerical simulations arises.

In comparison with the two-dimensional Bretherton flow (circular tubes, flow between parallel
plates) covered in our previous work \cite{kuzmin-binary2d}, there is a vast number of experimental
results available for the three-dimensional case as microchannels with square, triangular and
rectangular crosssections. For instance, \citet{cerro-bubble-train} performed a number of
experiments for a
bubble-train flow in capillaries of
circular and square cross section for horizontal, upward and downward flows. They found a
significant difference in mass transfer between upward and downward flows for the small capillary
number region.

\citet{shikazono-square} obtained
experimental
results for the deposition length dependency on the
capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
shaped mirochannels. They also found experimental correlation for the prediction of bubble radii
based on the capillary number $Ca$ and the Weber number $We=Re Ca$. Thus, there is the effect of
inertia on the bubble formation. It is indicated by \citet{gupta-review} that in comparison with
one phase flows the flows in microchannels depend on the complicated interplay of the gravitational,
interfacial, inertial and viscous forces.
{\color{red} Historically,  it was indicated \cite{bretherton} that inertia effects can be 
neglicted. Later on, the results of \citet{giavedoni-numerical} and
\citet{heil-bretherton} show
{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
range of Reynolds
numbers. 
For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
are negligible for $Ca\leq0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
number from $0$ to $70$ for the flow between plates. Later on, \citet{heil-bretherton} extended
results for the flow between plates upto Reynolds number
$300$. The indicate that while the influence of the established film thickness is insignificant
($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
significantly the pressure distrubution and the flow field near the front bubble tip. Recent
experimental results by \citet{shikazono-square} for square shaped capillaries indicate the
following dependency on the capillary number $Ca$ and the Weber number $We$:
\begin{equation}
\begin{aligned}
&R_{diag}=1.171-\frac{2.43 Ca^{2/3}}{1+7.28 Ca^{2/3}-0.255 We^{0.215}}\\
&R_{axis}=
\begin{cases}
1, &R_{diag}>1\\
R_{diag}, &R_{diag}\leq 1.
\end{cases}
\end{aligned}
\end{equation}
In the present
simulations the largest Reynolds number we achieve is less than $1$ and the capillary number is of
order of unity. Thus, it allows to neglect 
inertia effects. 
Therefore, the major governing parameter for microchannel Bretherton flows is not the density ratio,
but the
viscosity ratio. Thus, the applicability of the binary liquid model can be examined for the
simulation of the square shaped capillaries.

 Moreover, the results based on uniform density, as in the case of the LBM
binary liquid model, are in good agreement with other simulations
\cite{giavedoni-numerical,heil-bretherton} in which inertia effects were taken into account. The
same indicated in \cite{gupta-review} the  classical Bretherton problem is stated for
gas and liquid, which are of
significantly
different densities and viscosities.
}

The interested reader
is referred to these works for comprehensive experimental correlations.

The experimental works supported by numerical simulations reveal interesting phenomena happening in
three-dimensional geometry cases. It was found \cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for a
certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
symmetric case,
Fig. \ref{fig:sym:antisym}.
{\color{red} From hereon, we limit ourselves to the case of microchannels with square
crosssections.}
Non axisymmetric case is attributed to the case where the axial radius is different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. 
{\color{red}
\citet{cerro-space} indicate that because of asymmetry the flow in the corner can predominate the
flow in the side of the capillary thus changing the qualitative differencies depending on the
gravity direction. ``Because of the mixing, the gas-liquid mass tranfer coefficient are 2-5 times
larger fro upflow than for downflow''. 
}
In
this case
the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the bubble radii
dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.

\citet{heil-threedim} also indicated another interesting phenomenon. They indicated the transition
capillary number where the streamlines pattern
changes from having a vortex in front of the bubble and not having it for larger capilllary
numbers,
Fig. \ref{fig:streamlines:pattern}.
For the square channel the critical number is $Ca=0.691$.   
%\begin{figure}
%\includegraphics[width=0.5\textwidth]{Figures/ca05_clean.eps}\\
%\includegraphics[width=0.5\textwidth]{Figures/ca10_clean.eps}
%\caption{Two different regimes for recirculation. Courtesy of \citet{heil-threedim}. 
%Streamlines in the moving frame of reference for a tube with $Ca = 0.5$ (top) and $Ca = 2.0$
%(bottom). 
%\label{fig:heil:recirculating}}
%\end{figure}

There are also a number of numerical works for the three-dimensional case. For instance, 
\citet{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and calculated bubble shapes for different
slug and bubble cross sections and menisci appearance.
\citet{heil-threedim} performed three-dimensional simulations for circular shaped,
square and rectangular shaped capillaries. Also, they investigated gravitational effects for
circular shaped capillaries. They also found an empirical correlation which allows to put
different  aspect ratio rectangular channels simulation results on one
curve. \citet{wang-non-circular} performed numerical simulations by the VOF technique for the
non-circular
shaped capillaries as square and triangular shaped capillaries. The VOF technique is the continuous
interface method based on a non-uniform grid. It is of particular interest for us to compare it
with the continuous interface method by the LBM implementation. 
They also performed
numerical simulations for the determination of the critical capillary number $\widehat{Ca}$.

Almost all the simulations indicated above for the two-dimensional and three-dimensional cases
are performed with the help of the finite element method
\cite{ingham-plates,heil-bretherton,heil-threedim}. However, this method is of limited
applicability for
problems
involving complex geometries, free interface motion, or for problems involving coalescence and/or
droplet breakup. The continuous interface models are more flexible for this kind of simulations.
However, if the interface is smeared out over several grid nodes, the question of
proper film resolution in comparison with the interface resolution arises.  Therefore, the
\emph{goal} of this work is to reproduce the numerical results and phenomena presented above
with the help of the continuous interface binary liquid free energy lattice Boltzmann model. We
focus on exploring the parameter range of the binary liquid lattice
Boltzmann method for correctly resolving the flow in a range of capillary numbers.

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. Thanks to its kinetic nature, LBM as a particle
method easily tackles complex geometries and allows incorporation of
physical phenomena on the microscopic level, as in the case of multiphase models. Most
multiphase lattice Boltzmann models \cite{swift, Shan-chen:extended} resolve
the interface using continuous interface methods where the interface spans over several grid nodes.
Such a representation
brings issues of the film thickness resolution versus interface
resolution -- the diffuse interface should be dealt with in such a way as to have a
negligible effect on the physics of the film. In our previous work \citet{kuzmin-binary2d} the grid
resolution was examined. It was found that the film thickness should be at least as twice as the
interface for the simulation to be accurate.

The binary liquid free-energy LB model due to \citet{swift} we used
simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for gas and liquid, which are of
significantly
different densities and viscosities, it was indicated \cite{bretherton} that inertia effects can be 
neglected. Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
numbers.
Therefore, a major governing factor for microchannel flows is not the density ratio, but the
viscosity ratio. The goal of this work is to do a feasibility study of the LBM
binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
problem. The work results are in good agreement with other simulations
\cite{giavedoni-numerical,heil-bretherton,heil-threedim}.

One should also acknowledge the works of \citet{pagonabarraga-fingers} on menisci
in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
simulations of two-dimensional channel flows for a relatively large capillary numbers, and
found discrepancies with the classical Bretherton theory, which
is limited to the low capillary number regime \cite{giavedoni-numerical}.

The paper is organized as follows.  First, we briefly
explain the binary liquid lattice Boltzmann model. Then, the preliminary results for the
two-dimensional and for the three-dimensional case are presented in the results section. The paper
is
concluded with a summary of the main findings.
